The growth of children's semantic and phonological networks: insight from 10 languages

### 2019-09-23 journal/ms.Rmd
new version

What factors shape vocabulary learning over the course of early childhood? To investigate this question, scientists have adopted multiple research strategies, from conducting controlled laboratory experiments [e.g. @markman90] to analyzing dense corpora capturing language learning in context [e.g., @roy2015]. One prominent strategy consists in documenting the timeline of words' acquisition, and studying the properties that make words easy or hard to learn [e.g., @goodman2008; @huttenlocher1991]. For example, @goodman2008 found that, within a lexical category (e.g., nouns), higher parental frequency is associated with earlier learning. Researchers have studied the role of several other factors such as word length and the mean length of utterances in which the word occurs [e.g., @braginsky2019; @swingley2018].

Besides word-level properties, the structure of the lexicon (that is, how words relate to one another) is also linked to the the Age of Acquisition (AoA) of words. The lexical structure can be characterized in terms of a network where each node represents a word in the vocabulary, and each link between two nodes represents a relationship between the corresponding pair of words [e.g., @collins1975; @luce1998]. Previous studies have investigated early vocabulary structure by constructing networks using  a variety of word-word relations including shared semantic features [@mcrae2005], target-cue relationships in free association norms [@nelson1998], co-occurrence in child directed speech [@macwhinney2014], and phonological relatedness [@vitevitch2008]. These studies have generally found that children tend to produce words that have higher neighborhood density (i.e., high connectivity in the network) earlier, both at the phonological and the semantic level [@storkel2009; @hills2009; @hills2010; @stella2017; @carlson2014].

@@

<!-- For example, having a rich and organized knowledge about the domain of dinosaurs would facilitate the acquisition of new dinosaur-related words [@chi1983]. -->
These two proposals represent two divergent ideas about the role of lexical networks in acquisition. On the INT proposal, learning is driven by known words with high connectivity to other known words (Figure 1, left). Thus, the network structure is a causal factor in word learning, that is, children rely on the organization of their past knowledge to determine future learning [@storkel2009; @altvater2013; @chi1983; @borovsky2016]. In contrast, on the EXT approach, learning is driven by the connectivity of words that are not known yet (Figure 1, right). Thus, the relevant network structure is not internally represented by children, and the observed connectivity effect might be an epiphenomenon of some properties of the linguistic input. For example, highly connected concrete nouns in the input could be more easily learned because of their contextual diversity, allowing for an easier meaning disambiguation [@smith2008; @mcmurray2012; @yurovsky2015]. Another reason could be that these words are emphasized by the caregivers in their child-directed speech [@huttenlocher1991; @hoff2002; @clark2007].

```{r growth, fig.cap = "Illustration of the two growth scenarios. Filled grey circles represent known words (Internal) at a certain point in time. The empty, numbered circles represent words that have not yet been learned (External) and which are candidates to enter the lexicon next. The identity of the word that is going to be learned depends on the growth scenario.  Here the squares indicate the node that drives growth in each scenario and the hand pointer indicates which word is likely to be learned. For INT, the utility of a candidate, external node is the average degree (i.e., number of links) of the internal nodes that it would attach to. In this simplified example, candidate node 1 would connect to an internal node with 3 connections, thus we have $u_{INT}(node_1)= 3$. As for candidate node 2, it would connect to internal nodes that have only one connection each, making an average of 1, i.e., $u_{INT}(node_2)= 1$. According to INT, node 1 is more likely to enter the lexicon. For EXT, the utility of a candidate node is its degree in the entire network. In our example, candidate node 1 has 2 connections in total whereas candidate node 2 has 5 connections. So we have $u_{EXT}(node_1)= 2$ and $u_{EXT}(node_2)= 5$. Thus, according to EXT, node 2 is more likely to enter the lexicon next. This figure is based on an example from Hills et al. (2009).", fig.align = "center", out.width = "300px"}
knitr::include_graphics("figs/growth3.png")
```

@@

First, we study the phonological network in addition to the semantic network. These two networks represent different ways the mental lexicon is structured [@beckage2016]. In particular, words that are neighbors in the semantic network (e.g., "cat", "dog") are not necessarily neighbors in the phonological network, and vice versa. Does the phonological network also predict word learning? Previous work has found an effect of words' connectivity in the phonological network on their age of learning [@storkel2009; @stella2017; @carlson2014]. In other words, words learned earlier in life tend to sound similar to many other words than a word learned later in life. However, this finding is *a priori* compatible with both INT and EXT, and previous studies did not explicitly compare these two mechanisms.  Here, we investigate whether phonological networks, like semantic networks, grow through EXT, or if they rather grow via INT (Figure\ \@ref(fig:growth)).
<!--
The phonological strcuture influence word processing in adults (Vitevitch, 2008;  Chan & Vitevitch, 2009; Chan & Vitevitch,2010; Vitevitch, Chan, and Roodenrys, 2012; Castro et al. 2018). Here we test whether the mechanism of networtk growth in the phonological domains mirrors the mechanism found in the semantic domains. -->

@@

We adopted the following research strategy. We used parent reports on the MacArthur-Bates Communicative Development Inventory and its cross-linguistic adaptations [@frank2017; @fenson94]. We studied the timeline of word learning using the normative age of acquisition (i.e., the age at which at least 50% of children know a given words). Our choice of studying the normative learning trajectory rather than the individual trajectories was motivated by the nature of the dataset used---which is primarily based on cross-sectional studies. Children may vary in their individual learning trajectories, but the aggregate data provide highly robust measures of the average learning patterns [@fenson94]. The use of such measures has lead to important insights on the mechanisms of word learning [@storkel2009; @stella2017; @hills2009; @hills2010; @goodman2008].

The paper is organized as follows. First, we describe the datasets we used and explain how we constructed the networks. Second, we analyze static properties of words' connectivity in these networks (correlation with age of acquisition and shape of the distribution) and we explain how these properties inform hypotheses about network growth. Next, we fit the two hypothesized growth mechanisms to the data. We investigate the extent to which the results obtained in @hills2009 generalize to phonological networks and comprehension-based vocabularies, and whether this generalization holds cross-linguistically.

@@

We constructed semantic networks for English data following the procedure outlined in @hills2009, as follows. We used as an index of semantic relatedness the Florida Free Association Norms [@nelson1998]. This dataset was collected by giving adult participants a word (the cue), and asking them to write the first word that comes to mind (the target). For example, when given the word "ball", they might answer with the word "game". A pair of nodes were connected by a directed link from the cue to the target if there was a cue-target relationship between these nodes in the association norms. The connectivity of a given node was characterized by its *indegree*: the number of links for which the word was the target.\footnote{This choice was based on prior work by Hills et al. (2009) stating that analyses with both outdegrees (sum of the links where the word is the cue in a cue-target pair) and total degree  (outdegree plus indegree) led to results weaker than those calculated with indegree.} To model growth from month to month, we constructed a different network at each month, based on the nouns that have been acquired by that month.

Since the free association norms are available only in English, we used the hand-checked translation equivalents available in Wordbank, which allowed us to use the English association norms across 10 languages. Semantic associations are not necessarily shared across languages, but we use this technique as a reasonable first approximation. In support of this approximation, @youn2016 showed that semantic networks across languages share substantial similarities. Furthermore, using the same association data across languages does not necessarily mean that the resulting networks will grow in a similar fashion. For instance, the set of words acquired by children as well as the order of word acquisition can vary from language to language leading to possibility different learning strategies.

@@

To construct phonological networks we first mapped the orthographic transcription of words to their International Phonetic Alphabet (IPA) transcriptions in each language, using the open source text-to-speech software **[Espeak](http://http://espeak.sourceforge.net/)**. This software provides the correct IPA transcription if the word is found in a spelling-to-phonemes dictionary, otherwise it uses  language-specific pronunciation rules to generate an approximate phonetic transcription. We used the Levenshtein distance (also known as edit distance) as a measure of phonological relatedness between two nodes. The measure counts the minimum number of operations (insertions, deletions, substitutions) required to change one string into another.

In previous studies, two nodes were linked if they had an edit distance of 1 [@storkel2009; @carlson2014; @stella2017]. However, these studies reported a contribution of phonological connectivity to noun learning when networks were built using a rich adult vocabulary. Since the focus of the current study is on the mechanism of growth, the networks should be based on children's early vocabulary. The latter, however, contains very few noun pairs with an edit distance of 1. Thus, we increased the threshold from 1 to 2, that is, two nodes were related if their edit distance was equal to 1 or 2.\footnote{We also considered the case of an edit distance of 1 as well as the continuous measure, i.e., the inverse edit distance without threshold. In both cases, the results were weaker than those obtained with a threshold of 2. We did not consider the case of a threshold larger than 2 since many short pairs appear phonologically unrelated when the edit distance is 3 or more (e.g., "cat"/"dog").} The connectivity of a given node was characterized with its *degree*: the number of links it shares with other words.

@@

In brief, the analysis consisted in two steps. First, we derived the optimal cut-off, $k_{min}$, above which the distribution is more likely to follow a power law,\footnote{In natural phenomena, it is often the case that the power law applies only for values above a certain minimum.} and we estimate the corresponding scaling parameter $\alpha$. Second we calculated the goodness-to-fit, which resulted in a $p$-value quantifying the plausibility of the model. The results are shown in Table \@ref(tab:powerLawProd) for production data, and in Table \@ref(tab:powerLawComp) for comprehension data.

@@

\noindent where $\beta$ is a fitted parameter that captures the magnitude of the relationship between network parameters and growth (analogous to a regression coefficient). A positive value of $\beta$ means that words with higher utility values $u_i$ are acquired first, and a negative value means that words with lower utility values are acquired first (see Figure \@ref(fig:growth) for an illustration of how utilities values $u_i$ are defined in each growth scenario). The normalization includes all words that could be learned at that month.

We estimated the parameter $\beta$ using a Bayesian approach. The inference was performed using the probabilistic programming language  WebPPL [@dippl]. We defined a uniform prior over $\beta$, and at each month, we computed the likelihood function over words that could possibly enter the lexicon at that month, fit to the words that have been learned at that month (using Formula 1). Markov Chain Monte Carlo sampling resulted in a posterior distribution over $\beta$, which we summarized in Figure \@ref(fig:growthPred). The results replicate Hills et al.'s original finding regarding the semantic network in English and the production-based vocabulary, which is that this network grows by EXT, not by INT. Crucially, our results show that, generally speaking, this finding generalizes to comprehension-based vocabulary, and holds across languages. This generalization was obtained in both the semantic^[One could imagine that the fact of using English free association norms cross-linguistically would decrease the effect of non-English semantic networks because of possible cultural differences. However, our findings do not support this assumption, rather, it supports our initial approximation about the shared nature of the semantic similarity measure. That said, this approximation is not perfect. For example there is evidence that a small part of the variance in free association data can be explained by phonological similarity [@kachergis2011; @matusevych2018], thus leading to possibly minor cross-linguistic differences.] and phonological domains.

@@

Above we showed that the way semantic and phonological information is structured in the learning environment contributes to noun learning (via EXT) across languages. However, we know that other factors influence learning as well [e.g., @braginsky2019]. Next we investigated how semantic and phonological connectivity interact with two other factors. The first one is word frequency, a well studied factor shown to predict the age of acquisition in a reliable fashion [e.g. @goodman2008]. The second factor is word length, which was shown to correlate with phonological connectivity: Shorter words are more likely to have higher connectivity [@pisoni1985; @vitevitch2005].

Since we found INT to be uninformative, we dropped it from this analysis, keeping only EXT. This simplified the model because we no longer needed to fit growth month-by-month. The latter was a requirement only for INT where the words' utilities varied from month to month, depending on how connectivity changed in the growing internal network. A more direct way to assess and compare the contribution of EXT in relation to other word-level factors is through conducting regressions, where connectivity in the learning environment, frequency and length predict the age of acquisition.

@@

For example, in the semantic domain, growth according to EXT could be explained by a mechanism similar to cross-situational learning [@smith2008; @mcmurray2012; @yurovsky2015]. According to this mechanism, children track the co-occurrence of concrete nouns with their possible semantic referents. The referent of a word heard in only one naming situation can be ambiguous (e.g., when the word "ball" is heard for the first time in the presence of both a ball and a chair), but hearing the same word in a diversity of semantic contexts allows the learner to narrow down the set of possible word-object mappings. In our case, free association (used to determine semantic network connectivity) is related to contextual co-occurrence [@griffiths07; @fourtassi2013], meaning that highly connected words will tend to occur in a variety of speech and referential contexts.  This fact makes such words easier to learn because they have more referential disambiguating cues across learning contexts. Crucially, children can learn these words without necessarily knowing the meaning of all other words with which they co-occur (hence the similarity with EXT). This possibility is supported by the finding that words' diversity of occurrence in child directed speech predicts their age of learning [@stella2017; @hills2010].

In the phonological case, network growth according to EXT is also compatible with a scenario whereby children are tracking low level statistical patterns, e.g., high probability sound sequences. Indeed, connectivity in the phonological network is inherently correlated with phonotactic probability [@vitevitch1999]. That is, highly connected words tend to be made of frequent sound sequences. Children are sensitive to local phonotactic regularities [@jusczyk1994] and this sensitivity might lead them to learn higher-probability words more easily [@storkel2001]. This explanation is supported by computational simulations that show how learning general phonotactics patterns create "well-worn paths" which allow the models to represent several distinct but phonologically neighboring words [@dell1993; @takac2017; @siew2013]. More generally, there is a growing interest in investigating precisely how the local patterns acquired through statistical learning may give rise to the global network orgnization [For a review, see @karuza2016].

Besides using their own statistical learning skills, children could also benefit from the way their caregivers speak. Perhaps the caregivers put more emphasis on the words that are highly connected in *their* mature lexical network. This emphasis would guide children to learn first these highly connected words even though children do not have access to the distribution of words' connectivity in the final network.  Investigating this possibility would require further research on caregiver-child interaction [@macwhinney2014; @roy2015], examining what words are introduced over development and the extent to which children's uptake is influenced by this input [@huttenlocher1991; @clark2007; @hoff2002].

This study investigated the class of nouns in isolation --- following previous studies investigating the early semantic and phonological network [@hills2009; @storkel2009]. We could ask if studying one class separtaly is a legitimate research strategy. In other words, would word classes (such as nouns, verbs and function words) be acquired relatively differently, or would they interact substantially to the extent that it becomes unreasonable to study each class separately?

We know that different word classes follow different time trajectories. For instance, in the early stages of development, nouns tend to be acquired at a higher pace than predicates and function words [@bates1994]. Research has shown that this difference cannot be trivially attributed to differences in the degree to which these classes are present in the input; if anything, verbs and function words are often more frequent in the input than nouns [e.g., @gentner1982].  @goodman2008 found an effect of frequency on the age of acquisition within --- not across --- classes. Further, a recent work by @braginsky2019 tested a large number of predictors, besides frequency, and found that these predictors do not influence the acquisition of word classes in the same way. For example, the acquisition of nouns was found to be most influenced by frequency and concreteness, whereas the acquisition of function words was most influenced by word length.

These patterns support the hypothesis that different word classes have different pathways of learning, thus making it worthwhile to study each class separately. This hypothesis, however, does not preclude all sorts of inter-class interaction. For example, @dautriche2015 found that the acquisition of similar-sounding words ("bih"-"dih") is easier if these words belong to different syntactic categories (e.g., a noun and a verb).

This work shares a number of limitations with previous studies using similar research strategy and datasets. Chief among these limitations is the fact that the age of word acquisition is computed using different children at different ages (due to the fact that available CDI data is mainly cross-sectional). Such a measure has been shown to be valid and reliable [@fenson94], and has allowed researchers to study important aspects of word learning [@hills2009; @stella2017; @storkel2009; @goodman2008; @braginsky2019]. In our case, the use of cost-effective cross-sectional data has allowed us to leverage large-scale studies across several languages. That said, it is important to remember that this type of data can only inform us about the learning trajectory of the "average" child. Although our study endorses, overall, the externally-driven account of network growth, this does not mean individual children never use some variant of INT or some combination of both INT and EXT [@beckage]. To illustrate, some children develop "islands of expertise", that is, well organized knowledge about a certain topic (e.g., birds or dinosaurs). This prior knowledge enables these children to learn new related words more easily [e.g., @chi1983].

@@

That said, it is useful to include the results obtained with both thresholds. In addition, it could be useful to compare the results to the case of weighted networks, i.e., with no thresholding. The main analyses for these two cases are shown in what follows. In summary, both methods of constructing the phonological networks had issues. While the fact of using edit = 1 created sparsity, using weighted networks created colinearity with length.

@@

We show in Figure \@ref(fig:corrPlot) the correlation between the phonological connectivity and age of acquisition in both comprehension and production. The sparsity issue, due the low phonological neighborhood in the children's lexicon is apparent: most words had 0 degree and a few had non-zero but small degrees. The values of the correlations are much lower than the ones obtained with the threshold of 2.

The next figures show how the phonology-based mechanism of growth (phonoEXT) fares in comaprision to semEXT and other predictors of learing in each language (Figure \@ref(fig:staticPredEdit1) and across all languages (Figure \@ref(fig:staticAllEdit)). These figures show that phonoEXT based on edit distance 1 had no noticeable effect on learning.

@@

```{r corrPlot, out.width = "\\textwidth", fig.pos = "!h", fig.height=5, fig.cap = "Age of acquistion in both comprehension and production as predicted by the degree (i.e., connectivity) in phonological network, using an edit distance of 1. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients."}

@@

```{r staticPredEdit1, out.width = "\\textwidth", fig.pos = "!h", fig.cap = "Estimates of the relative contribution of each predictor of AoA in the regression model in each language. Results are shown for both production and comprehension data. Ranges indicate 95\\% confidence intervals. Positive values indicate a positive relationship (e.g. longer words tend to have a higher AoA), while negative values indicate a negative relationship (e.g. words with higher frequency tend to have a lower AoA).", fig.width=10, fig.height=4}

@@

```{r staticAllEdit, out.width = "\\textwidth", fig.pos = "!h", fig.cap = "Estimates of the relative contribution of each predictor of AoA in the combined model. Here the phonologocal networks were based on an edit distance of 1. Results are shown for both production and comprehension data. Ranges indicate 95\\% confidence intervals. Dotted ranges indicate the estimates for the predictor in a separate model that includes only this predictor as a fixed effect.", fig.width=6, fig.height=2.5}

@@

## Analyses using wighted phonological networks with no thresholding

>>>>

What factors shape vocabulary learning over the course of early childhood? To investigate this question, scientists have adopted multiple research strategies, from conducting controlled laboratory experiments [e.g. @markman90] to analyzing dense corpora capturing language learning in context [e.g., @roy2015]. One prominent strategy consists in documenting the timeline of words' acquisition and studying the properties that make words easy or hard to learn [e.g., @goodman2008; @huttenlocher1991]. For example, @goodman2008 found that, within a lexical category (e.g., nouns), higher parental frequency is associated with earlier learning. Researchers have studied the role of several other factors such as word length and the mean length of utterances in which the word occurs [e.g., @braginsky2019; @swingley2018].

Besides word-level properties, the structure of the lexicon (that is, how words relate to one another) is also linked to the Age of Acquisition (AoA) of words. The lexical structure can be characterized in terms of a network where each node represents a word in the vocabulary, and each link between two nodes represents a relationship between the corresponding pair of words [e.g., @collins1975; @luce1998]. Previous studies have investigated early vocabulary structure by constructing networks using  a variety of word-word relations including shared semantic features [@mcrae2005], target-cue relationships in free association norms [@nelson1998], co-occurrence in child-directed speech [@macwhinney2014], and phonological relatedness [@vitevitch2008]. These studies have generally found that children tend to produce words that have higher neighborhood density (i.e., high connectivity in the network) earlier, both at the phonological and the semantic level [@storkel2009; @hills2009; @hills2010; @stella2017; @carlson2014].

@@

<!-- For example, having a rich and organized knowledge about the domain of dinosaurs would facilitate the acquisition of new dinosaur-related words [@chi1983]. -->
These two proposals represent two divergent ideas about the role of lexical networks in acquisition. On the INT proposal, learning is driven by known words with high connectivity to other known words (Figure 1, left). Thus, the network structure is a causal factor in word learning, that is, children rely on the organization of their past knowledge to determine future learning [@storkel2009; @altvater2013; @chi1983; @borovsky2016]. In contrast, on the EXT approach, learning is driven by the connectivity of words that are not known yet (Figure 1, right). Thus, the relevant network structure is not internally represented by children, and the observed connectivity effect might be an epiphenomenon of some properties of the linguistic input. For example, highly connected concrete nouns in the input could be more easily learned because of their contextual diversity, allowing for easier meaning disambiguation [@smith2008; @mcmurray2012; @yurovsky2015]. Another reason could be that these words are emphasized by the caregivers in their child-directed speech [@huttenlocher1991; @hoff2002; @clark2007].

```{r growth, fig.cap = "Illustration of the two growth scenarios. Filled grey circles represent known words (Internal) at a certain point in time. The empty, numbered circles represent words that have not yet been learned (External) and which are candidates to enter the lexicon next. The identity of the word that is going to be learned depends on the growth scenario.  Here the squares indicate the node that drives growth in each scenario, and the hand pointer indicates which word is likely to be learned. For INT, the utility of a candidate external node is the average degree (i.e., number of links) of the internal nodes that it would attach to. In this simplified example, candidate node 1 would connect to an internal node with 3 connections; thus we have $u_{INT}(node_1)= 3$. As for candidate node 2, it would connect to internal nodes that have only one connection each, making an average of 1, i.e., $u_{INT}(node_2)= 1$. According to INT, node 1 is more likely to enter the lexicon. For EXT, the utility of a candidate node is its degree in the entire network. In our example, candidate node 1 has 2 connections in total, whereas candidate node 2 has 5 connections. So we have $u_{EXT}(node_1)= 2$ and $u_{EXT}(node_2)= 5$. Thus, according to EXT, node 2 is more likely to enter the lexicon next. This figure is based on an example from Hills et al. (2009).", fig.align = "center", out.width = "300px"}
knitr::include_graphics("figs/growth3.png")
```

@@

First, we study the phonological network in addition to the semantic network. These two networks represent different ways the mental lexicon is structured [@beckage2016]. In particular, words that are neighbors in the semantic network (e.g., cat, dog) are not necessarily neighbors in the phonological network and vice versa. Does the phonological network also predict word learning? Previous work has found an effect of words' connectivity in the phonological network on their age of learning [@storkel2009; @stella2017; @carlson2014]. In other words, words learned earlier in life tend to sound similar to many other words than a word learned later in life. However, this finding is *a priori* compatible with both INT and EXT, and previous studies did not explicitly compare these two mechanisms.  Here, we investigate whether phonological networks, like semantic networks, grow through EXT, or if they rather grow via INT (Figure\ \@ref(fig:growth)).
<!--
The phonological strcuture influence word processing in adults (Vitevitch, 2008;  Chan & Vitevitch, 2009; Chan & Vitevitch,2010; Vitevitch, Chan, and Roodenrys, 2012; Castro et al. 2018). Here we test whether the mechanism of networtk growth in the phonological domains mirrors the mechanism found in the semantic domains. -->

@@

We adopted the following research strategy. We used parent reports on the MacArthur-Bates Communicative Development Inventory and its cross-linguistic adaptations [@frank2017; @fenson94]. We studied the timeline of word learning using the normative age of acquisition (i.e., the age at which at least 50% of children know a given word). Our choice of studying the normative learning trajectory rather than the individual trajectories was motivated by the nature of the dataset used---which is primarily based on cross-sectional studies. Children may vary in their individual learning trajectories, but the aggregate data provide highly robust measures of the average learning patterns [@fenson94]. The use of such measures has lead to important insights on the mechanisms of word learning [@storkel2009; @stella2017; @hills2009; @hills2010; @goodman2008].

The paper is organized as follows. First, we describe the datasets we used and explain how we constructed the networks. Second, we analyze static properties of words' connectivity in these networks (correlation with age of acquisition and shape of the distribution), and we explain how these properties inform hypotheses about network growth. Next, we fit the two hypothesized growth mechanisms to the data. We investigate the extent to which the results obtained in @hills2009 generalize to phonological networks and comprehension-based vocabularies, and whether this generalization holds cross-linguistically.

@@

We constructed semantic networks for English data following the procedure outlined in @hills2009, as follows. We used as an index of semantic relatedness the Florida Free Association Norms [@nelson1998]. This dataset was collected by giving adult participants a word (the cue), and asking them to write the first word that comes to mind (the target). For example, when given the word "ball," they might answer with the word "game." A pair of nodes were connected by a directed link from the cue to the target if there was a cue-target relationship between these nodes in the association norms. The connectivity of a given node was characterized by its *indegree*: the number of links for which the word was the target.\footnote{This choice was based on prior work by Hills et al. (2009) stating that analyses with both outdegrees (sum of the links where the word is the cue in a cue-target pair) and total degree  (outdegree plus indegree) led to results weaker than those calculated with indegree.} To model growth from month to month, we constructed a different network at each month, based on the nouns that have been acquired by that month.

Since the free association norms are available only in English, we used the hand-checked translation equivalents available in Wordbank, which allowed us to use the English association norms across 10 languages. Semantic associations are not necessarily shared across languages, but we use this technique as a reasonable first approximation. In support of this approximation, @youn2016 showed that semantic networks across languages share substantial similarities. The fact that semantic associations are assumed to be shared across languages does not mean that the semantic networks will necessarily grow in a similar fashion. For instance, the set of words acquired by children as well as the order of word acquisition can vary from language to language leading to possibily different learning strategies.

@@

To construct phonological networks we first mapped the orthographic transcription of words to their International Phonetic Alphabet (IPA) transcriptions in each language, using the open-source text-to-speech software **[Espeak](http://http://espeak.sourceforge.net/)**. This software provides the correct IPA transcription if the word is found in a spelling-to-phonemes dictionary, otherwise it uses  language-specific pronunciation rules to generate an approximate phonetic transcription. We used the Levenshtein distance (also known as edit distance) as a measure of phonological relatedness between two nodes. The measure counts the minimum number of operations (insertions, deletions, substitutions) required to change one string into another.

In previous studies, two nodes were linked if they had an edit distance of 1 [@storkel2009; @carlson2014; @stella2017]. However, these studies reported a contribution of phonological connectivity to noun learning when networks were built using a dense adult vocabulary. Since the focus of the current study is on the mechanism of growth, the networks are based on children's early vocabulary. The latter, however, contains very few noun pairs with an edit distance of 1. To better represent the similarity space in the phonological domain, we increased the threshold from 1 to 2, that is, two nodes were related if their edit distance was equal to 1 or 2.\footnote{In Appendix A, we show the main results for phonological networks based on an edit distance of 1. We also show the results for phonological networks where the edges between pairs of words were weighted by the inverse of the edit distance. We did not consider the case of a threshold larger than 2 since many short pairs appear phonologically unrelated when the edit distance is 3 or more (e.g., "cat"/"dog").} The connectivity of a given node was characterized with its *degree*: the number of links it shares with other words.

@@

In brief, the analysis consisted in two steps. First, we derived the optimal cut-off, $k_{min}$, above which the distribution is more likely to follow a power law,\footnote{In natural phenomena, it is often the case that the power law applies only for values above a certain minimum.} and we estimate the corresponding scaling parameter $\alpha$. Second, we calculated the goodness-to-fit, which resulted in a $p$-value quantifying the plausibility of the model. The results are shown in Table \@ref(tab:powerLawProd) for production data, and in Table \@ref(tab:powerLawComp) for comprehension data.

@@

\noindent where $\beta$ is a fitted parameter that captures the magnitude of the relationship between network parameters and growth (analogous to a regression coefficient). A positive value of $\beta$ means that words with higher utility values $u_i$ are acquired first, and a negative value means that words with lower utility values are acquired first (see Figure \@ref(fig:growth) for an illustration of how utility values $u_i$ are defined in each growth scenario). The normalization includes all words that could be learned at that month.

We estimated the parameter $\beta$ using a Bayesian approach. The inference was performed using the probabilistic programming language  WebPPL [@dippl]. We defined a uniform prior over $\beta$, and at each month, we computed the likelihood function over words that could possibly enter the lexicon at that month, fit to the words that have been learned at that month (using Formula 1). Markov Chain Monte Carlo sampling resulted in a posterior distribution over $\beta$, which we summarized in Figure \@ref(fig:growthPred). The results replicate Hills et al.'s original finding regarding the semantic network in English and the production-based vocabulary, which is that this network grows by EXT, not by INT. Crucially, our results show that, generally speaking, this finding generalizes to comprehension-based vocabulary, and holds across languages. This generalization was obtained in both the semantic^[One could imagine that the fact of using English free association norms cross-linguistically would decrease the effect of non-English semantic networks because of possible cultural differences. However, our findings do not support this assumption; rather, it supports our initial approximation about the shared nature of the semantic similarity measure. That said, this approximation is not perfect. For example, there is evidence that a small part of the variance in free association data can be explained by phonological similarity [@kachergis2011; @matusevych2018], thus leading to possibly minor cross-linguistic differences.] and phonological domains.

@@

Above we showed that the way semantic and phonological information is structured in the learning environment contributes to noun learning (via EXT) across languages. However, we know that other factors influence learning as well [e.g., @braginsky2019]. Next, we investigated how semantic and phonological connectivity interact with two other factors. The first one is word frequency, a well-studied factor shown to predict the age of acquisition in a reliable fashion [e.g., @goodman2008]. The second factor is word length, which was shown to correlate with phonological connectivity: Shorter words are more likely to have higher connectivity [@pisoni1985; @vitevitch2005].

Since we found INT to be uninformative, we dropped it from this analysis, keeping only EXT. This simplified the model because we no longer needed to fit growth month-by-month. The latter was a requirement only for INT where the words' utilities varied from month to month, depending on how connectivity changed in the growing internal network. A more direct way to assess and compare the contribution of EXT in relation to other word-level factors is through conducting regressions, where connectivity in the learning environment, frequency, and length predict the age of acquisition.

@@

For example, in the semantic domain, growth according to EXT could be explained by a mechanism similar to cross-situational learning [@smith2008; @mcmurray2012; @yurovsky2015]. According to this mechanism, children track the co-occurrence of concrete nouns with their possible semantic referents. The referent of a word heard in only one naming situation can be ambiguous (e.g., when the word "ball" is heard for the first time in the presence of both a ball and a chair), but hearing the same word in a diversity of semantic contexts allows the learner to narrow down the set of possible word-object mappings. In our case, free association (used to determine semantic network connectivity) is related to contextual co-occurrence [@griffiths07; @fourtassi2013], meaning that highly connected words will tend to occur in a variety of speech and referential contexts.  This fact makes such words easier to learn because they have more referential disambiguating cues across learning contexts. Crucially, children can learn these words without necessarily knowing the meaning of all other words with which they co-occur (hence the similarity with EXT). This possibility is supported by the finding that words' diversity of occurrence in child-directed speech predicts their age of learning [@stella2017; @hills2010].

In the phonological case, network growth according to EXT is also compatible with a scenario whereby children are tracking low-level statistical patterns, e.g., high probability sound sequences. Indeed, connectivity in the phonological network is inherently correlated with phonotactic probability [@vitevitch1999]. That is, highly connected words tend to be made of frequent sound sequences. Children are sensitive to local phonotactic regularities [@jusczyk1994], and this sensitivity might lead them to learn higher-probability words more easily [@storkel2001]. This explanation is supported by computational simulations that show how learning general phonotactics patterns create "well-worn paths" which allow the models to represent several distinct but phonologically neighboring words [@dell1993; @takac2017; @siew2013]. More generally, there is a growing interest in investigating precisely how the local patterns acquired through statistical learning may give rise to the global network organization [For a review, see @karuza2016].

Besides using their own statistical learning skills, children could also benefit from the way their caregivers speak. Perhaps the caregivers put more emphasis on the words that are highly connected in *their* mature lexical network. This emphasis would guide children to learn first these highly connected words, even though children do not have access to the distribution of words' connectivity in the final network.  Investigating this possibility would require further research on caregiver-child interaction [@macwhinney2014; @roy2015], examining what words are introduced over development and the extent to which children's uptake is influenced by this input [@huttenlocher1991; @clark2007; @hoff2002].

This study investigated the class of nouns in isolation --- following previous studies investigating the early semantic and phonological network [@hills2009; @storkel2009]. We could ask if studying one class separately is a legitimate research strategy. In other words, would word classes (such as nouns, verbs and function words) be acquired relatively differently, or would they interact substantially to the extent that it becomes unreasonable to study each class separately?

There are many observations that support the hypothesis that different word classes have different pathways of learning, making it worthwhile to study each class separately. For instance, different word classes follow different time trajectories: In the early stages of development, nouns tend to be acquired at a higher pace than predicates and function words [@bates1994]. Research has shown that this difference cannot be trivially attributed to differences in the degree to which these classes are present in the input; if anything, verbs and function words are often more frequent in the input than nouns [e.g., @gentner1982].  @goodman2008 found an effect of frequency on the age of acquisition within --- not across --- classes. Further, recent work by @braginsky2019 tested a large number of predictors, besides frequency, and found that these predictors do not influence the acquisition of word classes in the same way. For example, the acquisition of nouns was found to be most influenced by frequency and concreteness, whereas the acquisition of function words was most influenced by word length.

This work shares a number of limitations with previous studies using similar research strategy and datasets. Chief among these limitations is the fact that the age of word acquisition is computed using different children at different ages (due to the fact that available CDI data is mainly cross-sectional). Such a measure has been shown to be valid and reliable [@fenson94], and has allowed researchers to study important aspects of word learning [@hills2009; @stella2017; @storkel2009; @goodman2008; @braginsky2019]. In our case, the use of cost-effective cross-sectional data has allowed us to leverage large-scale studies across several languages. That said, it is important to remember that this type of data can only inform us about the learning trajectory of the "average" child. Although our study endorses, overall, the externally-driven account of network growth, this does not mean individual children never use some variant of INT or some combination of both INT and EXT [@beckage]. To illustrate, some children develop "islands of expertise," that is, well-organized knowledge about a certain topic (e.g., birds or dinosaurs). This prior knowledge enables these children to learn new related words more easily [e.g., @chi1983].

@@

That said, it is useful to include the results obtained with both thresholds. In addition, it could be useful to compare the results to the case of weighted networks, i.e., with no thresholding. The main analyses for these two cases are shown in what follows. In summary, both methods of constructing the phonological networks had issues. While the fact of using an edit distance of 1 created sparsity, using weighted networks created collinearity with length in the models.

@@

We show in Figure \@ref(fig:corrPlot) the correlation between the phonological connectivity and age of acquisition in both comprehension and production. The sparsity issue --- due to the low phonological neighborhood in the children's lexicon --- is apparent: Most words had 0 connectivity, and a few had non-zero but small degrees. The values of the correlations are much lower than the ones obtained with the threshold of 2.

The next figures show how the phonology-based mechanism of growth (phonoEXT) fares in comaprison to semEXT and other predictors of learning in each language (Figure \@ref(fig:staticPredEdit1) and across all languages (Figure \@ref(fig:staticAllEdit)). These figures show that phonoEXT based on edit distance 1 had no noticeable effect on learning.

@@

```{r corrPlot, out.width = "\\textwidth", fig.pos = "!h", fig.height=5, fig.cap = "Age of acquistion in both comprehension and production as predicted by the degree (i.e., connectivity) in the phonological networks, using an edit distance of 1. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients."}

@@

```{r staticPredEdit1, out.width = "\\textwidth", fig.pos = "!h", fig.cap = "Estimates of the relative contribution of each predictor of AoA in the regression models. The phonologocal networks were based on an edit distance of 1. Ranges indicate 95\\% confidence intervals. Positive values indicate a positive relationship (e.g. longer words tend to have a higher AoA), while negative values indicate a negative relationship (e.g. words with higher frequency tend to have a lower AoA).", fig.width=10, fig.height=4}

@@

```{r staticAllEdit, out.width = "\\textwidth", fig.pos = "!h", fig.cap = "Estimates of the relative contribution of each predictor of AoA in the combined model. The phonologocal networks were based on an edit distance of 1. Ranges indicate 95\\% confidence intervals. Dotted ranges indicate the estimates for the predictor in a separate model that includes only this predictor as a fixed effect.", fig.width=6, fig.height=2.5}

@@

## Analyses using weighted phonological networks with no thresholding
